Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices

### 2022-07-22 formalization.tex
[add] thread identifiers on figure 4
[fix] pfs -> pcs

In this work, we consider the following properties:
perfect forward secrecy, injective agreement for both the initiator and the
responder, and implicit agreement.
%
Agreement is considered on a set of parameters $S$ which also contains the
session key material \mSessKey{}.
%
We will first describe in detail both properties, and then describe the
contents of the set $S$.
%
We formalize these properties as shown in Figure~\ref{fig:props}, which is
a modified version of a figure in~\cite{Norr21}.
\begin{figure*}[ht]
\begin{align*}
    \mPredPfs \triangleq\ & \forall I, R, \mSessKey, t_2, t_3\mLogicDot
    \mK^{t_3}(\mSessKey)\  \land\ 
    (\mIComplete^{t_2}(I, R, \mSessKey)\, \lor\, \mRComplete^{t_2}(I, R, 
\mSessKey))
    \rightarrow\\
    &(\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)
    \ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)
    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(R, \mSessKey))
    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))
\end{align*}
\begin{align*}
    \mPredInjI \triangleq\ &
    \forall I, R, \mSessKey, S, t_2\mLogicDot \mIComplete^{t_2}(I, R, 
\mSessKey, S)
    \rightarrow\\
    &(\exists t_1\mLogicDot \mRStart^{t_1}(R, \mSessKey, S) \land t_1 \lessdot 
t_2)
    \land (\forall I' R' t_1' \mLogicDot \mIComplete^{t_1'}(I' , R', \mSessKey, S)
        \rightarrow t_1' \doteq t_1)
    \ \ \lor\ \ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)
\end{align*}
\begin{align*}
    \mPredInjR \triangleq\ &
    \forall I, R, \mSessKey, S, t_2\mLogicDot \mRComplete^{t_2}(I, R, 
\mSessKey, S)
    \rightarrow\\
    &(\exists t_1\mLogicDot \mIStart^{t_1}(I, R, \mSessKey, S) \land t_1 \lessdot 
t_2)
    \land (\forall I' R' t_1' \mLogicDot \mRComplete^{t_1'}(I' , R', \mSessKey, S)
        \rightarrow t_1' \doteq t_1)
    \ \ \lor\ \ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2).
\end{align*}
\begin{align*}
    \mPredImpI \triangleq\ &
    \forall I, R, \mSessKey, S, t_1\mLogicDot \mIComplete^{t_1}(I, R, \mSessKey, 
S)
    \rightarrow\\
      &(\forall I', R', S', t_2\mLogicDot \mRComplete^{t_2}(I', R', \mSessKey, S') 
\rightarrow
             (I=I' \land R=R' \land S=S'))\\
      &\land (\forall I', R', S', t_1'\mLogicDot
        (\mIComplete^{t_1'}(I', R', \mSessKey, S') \rightarrow t_1' \doteq t_1
        )\\
      &\ \ \ \lor(\exists t_0\mLogicDot \mRevLTK^{t_0}(R) \land t_0 \lessdot t_1)
    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(R, \mSessKey))
    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey)).
\end{align*}
%
\caption{Formalization of security properties and adversary model.}
\label{fig:props}
\end{figure*}

%----------------------------------------------------------------------- PFS
\subsubsection{Perfect Forward Secrecy (PFS).}
\label{sec:secrecy}
Informally, by perfect forward secrecy, we mean that the adversary cannot
gain access to the session key material \mSessKey{}, even if a long-term key is
leaked after the establishment of \mSessKey{}.
%
PFS for session key material \mSessKey{} is formalized as \mPredPfs{}
in Figure~\ref{fig:props}.
%

@@

The properties we listed in Section~\ref{sec:desired-properties}
translate directly into \mTamarin's logic.
%
We show the \mTamarin{} lemma which encodes the \mPredPfs{} property.
%
Other properties are formalized similarly. 
%

>>>>

In this work, we consider the following properties:
perfect forward secrecy, injective agreement for both the initiator and the
responder, and implicit agreement.
%
Agreement is considered on a set of parameters $S$ which also contains the
session key material \mSessKey{}.
%
We will first describe in detail both properties, and then describe the
contents of the set $S$.
%
We formalize these properties as shown in Figure~\ref{fig:props}, which is
a modified version of a figure in~\cite{Norr21}.
\begin{figure*}[ht]
\begin{align*}
    \mPredPcs \triangleq\ & \forall \mTID, I, R, \mSessKey, t_2, t_3\mLogicDot
    \mK^{t_3}(\mSessKey)\  \land\ 
    (\mIComplete^{t_2}(\mTID, I, R, \mSessKey)\, \lor\, \mRComplete^{t_2}(\mTID, I, R, 
\mSessKey))
    \rightarrow\\
    &(\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)
    \ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
    &\ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(R, \mSessKey))
    \ \lor\ (\exists t_1\mLogicDot \mRevEph^{t_1}(I, \mSessKey))\\
    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))\\
    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))\\[2em]
%
    \mPredInjI \triangleq\ &
    \forall \mTID_I, I, R, \mSessKey, S, t_2\mLogicDot \mIComplete^{t_2}(I, R, 
\mSessKey, S)
    \rightarrow\\
    &(\exists \mTID_R, t_1\mLogicDot \mRStart^{t_1}(\mTID_R, R, \mSessKey, S) \land t_1 \lessdot t_2
    \land (\forall \mTID_I', I', R', t_1' \mLogicDot \mIComplete^{t_1'}(\mTID_I', I' , R', \mSessKey, S)
        \rightarrow t_1' \doteq t_1))\\
    &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(R) \land t_1 \lessdot t_2)\\
    &\ \lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))\\[2em]
%
    \mPredInjR \triangleq\ &
    \forall \mTID_R, I, R, \mSessKey, S, t_2\mLogicDot \mRComplete^{t_2}(\mTID_R, I, R, 
\mSessKey, S)
    \rightarrow\\
    &(\exists \mTID_I, t_1\mLogicDot \mIStart^{t_1}(\mTID_I, I, R, \mSessKey, S) \land t_1 \lessdot t_2
    \land (\forall \mTID_R', I' R' t_1' \mLogicDot \mRComplete^{t_1'}(\mTID_R', I' , R', \mSessKey, S)
        \rightarrow t_1' \doteq t_1))\\
    &\ \lor\ (\exists t_1\mLogicDot \mRevLTK^{t_1}(I) \land t_1 \lessdot t_2)\\
    &\ \lor\ (\exists t_0, t_1\mLogicDot \mRStart^{t_0}(\mTID_R, R, \mSessKey, S) \land \mTEE^{t_1}(I) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))\\[2em]
%
    \mPredImpI \triangleq\ &
    \forall \mTID_I, I, R, \mSessKey, S, t_1\mLogicDot \mIComplete^{t_1}(\mTID_I, I, R, \mSessKey, 
S)
    \rightarrow\\
      &(\forall \mTID_R, I', R', S', t_2\mLogicDot \mRComplete^{t_2}(\mTID_R, I', R', \mSessKey, S') \rightarrow
             (I=I' \land R=R' \land S=S')\\
      &\ \ \land (\forall \mTID_I', I', R', S', t_1'\mLogicDot
        \mIComplete^{t_1'}(\mTID_I', I', R', \mSessKey, S') \rightarrow t_1' \doteq t_1
        ))\\
    &\lor(\exists t_0\mLogicDot \mRevLTK^{t_0}(R) \land t_0 \lessdot t_1)
    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(R, \mSessKey))
    \lor(\exists t_0\mLogicDot \mRevEph^{t_0}(I, \mSessKey))\\
    &\lor\ (\exists t_0, t_1\mLogicDot \mIStart^{t_0}(\mTID_I, I, R) \land \mTEE^{t_1}(R) \land (t_0 \lessdot t_1 \lor t_1 \lessdot t_2))
\end{align*}
%
\caption{Formalization of security properties and adversary model.}
\label{fig:props}
\end{figure*}

%----------------------------------------------------------------------- PFS
\subsubsection{Perfect Forward Secrecy (PFS).}
\label{sec:secrecy}
Informally, by perfect forward secrecy, we mean that the adversary cannot
gain access to the session key material \mSessKey{}, even if a long-term key is
leaked after the establishment of \mSessKey{}.
%
PFS for session key material \mSessKey{} is formalized as \mPredPcs{}
in Figure~\ref{fig:props}.
%

@@

The properties we listed in Section~\ref{sec:desired-properties}
translate directly into \mTamarin's logic.
%
We show the \mTamarin{} lemma which encodes the \mPredPcs{} property.
%
Other properties are formalized similarly. 
%
